Comparing R packages for calculation of persistent homology

### 2019-12-16 tdabench-paper/tdabench.Rmd
corrected format for CRAN packages

Introductory section which may include references in parentheses
\citep{Rlang}, or cite a reference such as \citet{Rlang} in the text.

@@

An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.
We use \texttt{readr} v1.3.1 to read rectangular data \cite{readr}, \texttt{ggplot2} v3.2.1 \citep{ggplot2} and \texttt{magick} v2.2 \citep{magick} to visualize data, \texttt{bench} v1.0.4 to collect benchmark data \citep{bench}, \texttt{TDA} v1.6.9 \citep{TDA} and \texttt{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \texttt{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the TDA and TDAstats R packages.
Within the TDA package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the TDAstats package, benchmark data was collected for the Ripser \citep{Ripser} library.
As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.
Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, TDAstats.
Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.

@@

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
\bibliography{tdabench}

>>>>

@@

<!--An exclusively R (v3.6.2) code pipeline was built for this project \citep{Rlang,ihaka1996}.-->
We use \CRANpkg{readr} v1.3.1 to read rectangular data \cite{readr}, \CRANpkg{ggplot2} v3.2.1 \citep{ggplot2} and \CRANpkg{magick} v2.2 \citep{magick} to visualize data, \CRANpkg{bench} v1.0.4 to collect benchmark data \citep{bench}, \CRANpkg{TDA} v1.6.9 \citep{TDA} and \CRANpkg{TDAstats} v0.4.1 \citep{TDAstats} to calculate persistent homology of Vietoris-Rips and alpha simplicial complices, and \CRANpkg{pryr} v0.1.4 for calculations involving R objects \citep{pryr}.
Runtime calculations were averaged over 10 iterations and are visualized as mean $\pm$ SD.
Datasets were generated by sampling functions in base R to generate points uniformly distributed over circles (dimension = 2), spheres (dimension = 3), filled squares (dimension = 2), filled cubes (dimension = 3, 4), and tori (dimension = 3).
The number of points per point cloud varied from 25 to 500 along intervals of 25 points, which were empiric limits chosen after pragmatic considerations of available computational resources.
For consistency, the minimum and maximum simplicial complex radii were predetermined for each point cloud and provided as parameters to the \CRANpkg{TDA} and \CRANpkg{TDAstats} R packages.
Within the \CRANpkg{TDA} package, benchmark data was collected for the GUDHI \citep{gudhi} and Dionysus \citep{Dionysus} libraries; within the \CRANpkg{TDAstats} package, benchmark data was collected for the Ripser \citep{Ripser} library.
As alpha complex calculation was only implemented in GUDHI, alpha complex benchmark data was naturally only collected for the single library.
Measuring memory usage proved challenging since all the libraries calculating persistent homology were implemented in another programming language and then wrapped in R.
Thus, memory burden was indirectly measured by using boundary matrix size as a proxy.
Given that Ripser optimizes persistent homology calculation by avoiding calculation of a boundary matrix, memory use benchmarks are not provided for Ripser and, consequently, \CRANpkg{TDAstats}.
Fully reproducible code for all numerical results and figures in this report can be found at https://github.com/eashwarsoma/TDA-benchmark.

@@

\bibliography{tdabench}
